---
title: "Using parallel.lsm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-parallel-lsm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(parallel.lsm)
library(landscapemetrics)
```

# What parallel functions are available?

The package provides a version of most functions available in *landscapemetrics*. Only a few functions such as those starting with "show\_" are explicitly excluded, as they likely are not useful in parallel. To see a full list of the functions created when loading *parallel.lsm*, use `list_parallel.lsm()`.

```{r}
list_parallel.lsm()
```

# How can I adapt my existing landscapemetrics code to this package?

*parallel.lsm* code is predicated on being able to split at least one argument into multiple parts, such as when you already plan to iterate over multiple rasters, sampling points, or metrics. Luckily, splitting the data is the most difficult part (and it can be quite confusing at times!). If you can split your data into separate elements, you only need to tell *parallel.lsm* what argument(s) you want to split between cores and change the function name by adding "paralllel." to the front.

Note that the number of splits you make will determine how many cores you use. For most computers, it is advised to use half of your total cores (if attempting to run things in the background or with very large rasters) or n - 1 cores (fast, but can hit a memory limit). Most often with raster data, the limiting factor is RAM rather than the number of cores.

------------------------------------------------------------------------

If you're doing something like this...

```{r}
data <- NULL
for (raster in rasters) {
  raster_data <- lsm_l_ai(landscape = raster)
  data <- rbind(data, raster_data)
}
```

Try this...

```{r}
n_cores <- 4
rasters <- unname(split(rasters, ceiling(seq_along(rasters) / n_cores)))
data <- parallel.lsm_l_ai(landscape = rasters, split_on = "landscape")
```

------------------------------------------------------------------------

If you're doing something like this...

```{r}
cohesion <- lsm_l_cohesion(landscape = raster)
iji <- lsm_l_iji(landscape = raster)
gyrate <- lsm_l_gyrate_mn(landscape = raster)
pd <- lsm_c_pd(landscape = raster)
```

Or something like this (but it's being really slow)...

```{r}
metrics <- c("lsm_l_cohesion", "lsm_l_iji", "lsm_l_gyrate_mn", "lsm_c_pd")
data <- calculate_lsm(landscape = raster, what = metrics)
```

Try this...

```{r}
metrics <- c("lsm_l_cohesion", "lsm_l_iji", "lsm_l_gyrate_mn", "lsm_c_pd")
parallel.calculate_lsm(landscape = raster, what = metrics, split_on = "what")
```

# Additional Tips

-   *terra* is recommended by the developers of *landscapemetrics* over the *raster* package, as things are now optimized for SpatRaster objects.

-   The `join_tbl` argument is included in all parallelized functions and allows you to return a list of tibbles rather than one large merged on. If you're original script does something similar rather than binding rows, use this to avoid a large rewrite.

-   `terra::crop()` can be a good tool if you run into issues with RAM usage while using sampling points. Instead of providing one raster covering the entire study area, split the raster into smaller sections covering a specific split subset of sampling points. In cases where you cannot easily split these points based on information such as what site they belong to, consider k-means clustering if needed.

-   Splitting on multiple arguments is possible by providing a vector with multiple argument names to `split_on`. The main utility of this is providing plot_id values, but this can also be used if you have settings such as the directions that should be different between different rasters or metrics.
